Bayesian Nonparametric System Reliability\\ using Sets of Priors

### 2015-12-16 ijar-npb-ms.tex
writing out parameter choice guidelines

\section{Nonparametric Bayesian Approach for Component Reliability}

@@

While for an informative prior the choice of $\nktzl$ is relevant for posterior inferences in case of prior-data conflict,
in a near-noninformative setting, the choice of $\nktzl$ is not relevant,
as both $\yktnl$ and $\yktnu$ are obtained with $\nktzu$;
indeed $\yktzl > 0$ and $\yktzu < 1$ can be chosen such that
$\frac{s^k_t}{n_k} \in [\yktzl, \yktzu]$ for all $t \in \big(\min(\vec{t}^k), \max(\vec{t}^k)\big)$.
Obviously, one cannot have prior-data conflict in case of prior near ignorance.

@@

To obtain the lower and upper bound for the system reliability function $\Rsys(t)$,
we now need to minimise and maximise Equation~\eqref{} over $\PtZi{1}, \ldots, \PtZi{K}$ for each $t$,
where the posterior predictive probabilities for $C^k_t$ are given by the beta-binomial pmf \eqref{eq:postpredCny}.
It turns out that we need to optimise over $\nktz$ ($k=1,\ldots,K$) only,
as lower and upper bounds for $\Rsys(t)$ are obtained for $\yktzl$ and $\yktzu$, respectively.

To see this,
remember that $\yktz$ gives the prior expected functioning probability for a component of type $k$ at time $t$.
From Equation~\eqref{eq:nyupdate} (right) we see that $\yktn$ is increasing in $\yktz$,
%the posterior parameter set bounds are the transformed prior parameter set bounds,
%i.e., $\yktzu$ corresponds to the upper bound of $\PktN$, for all $\nktz \in \PktZ$,
such that for each fixed $\nktz$, $\yktzu$ will maximise the posterior expected functioning probability $\yktn$.
%
Now, a higher posterior expected functioning probability means
higher probability for many components out of $m_k$ surviving,
i.e., probabilities for the larger $l_k$'s are higher, and in turn for the lower $l_k$'s are lower.
***Indeed, the Beta-binomial distribution is stochastically ordered for varying $\yktn$, i.e.,
its cmf \eqref{eq:postpredCnycmf} is decreasing in $\yktn$ for all $l_k$,
such that the cmfs for different values of $\yktn$ do not cross each other.***
This means that for each fixed $\nktz$, the Beta-binomial cmf \eqref{eq:postpredCnycmf} is minimized for $\yktzu$ in all $l_k$,
such that the probability weights given to high values of $l_k$ are maximised. %for $\yktzu$.
For coherent systems, $\Phi$ is non-decreasing in each of its arguments $l_1,\ldots,l_K$,
thus $\Rsys(t)$ is maximised by putting most probability weight on the highest $l_k$ values for each $k$,
which is in turn obtained for $\yktzu$, $k=1,\ldots,K$.
We therefore have
%
%for a certain $k$, a high probability on high $l_k$'s is never worse than a high probability on lower $l_k$'s,
%keeping all other component type expressions fixed.
%***Monotonicity in $\yz_{k,t}$, so we get, for each $t \in {\cal T}$,
\begin{align*}
\lefteqn{\lRsys(t \mid \vec{t}^1, \ldots, \vec{t}^K)}\\
 &= \min_{\PtZi{1}, \ldots, \PtZi{K}} \Rsys(t \mid \PtZi{1}, \ldots, \PtZi{K}, \vec{t}^1, \ldots, \vec{t}^K) \\
 &= \min_{\nz_{1,t},\ldots,\nz_{K,t}} 
    \sum_{l_1=0}^{m_1} \cdots \sum_{l_K=0}^{m_K} \Phi(l_1, \ldots, l_K)
                                                 \prod_{k=1}^K P(C^k_t = l_k \mid \yktzl, \nktz, s^k_t) \\
 &= \min_{\nz_{1,t},\ldots,\nz_{K,t}} 
    \sum_{l_1=0}^{m_1} \cdots \sum_{l_K=0}^{m_K} \Phi(l_1, \ldots, l_K) \times \\ & \hspace*{12ex}
    \prod_{k=1}^K {m_k \choose l_k} \frac{B(l_k + \nn_{k,t}\ynl_{k,t}, m_k - l_k + \nn_{k,t}(1-\ynl_{k,t}))}
                                         {B(\nn_{k,t}\ynl_{k,t}, \nn_{k,t}(1-\ynl_{k,t}))} \,.
\end{align*}

***How to choose the sets of priors for the survival function of type $k$ components,
i.e. how to choose $\PkZi{1}, \ldots \PkZi{\tmax}$?***

***not necessary to ensure $p^k_{t_j} \ge p^k_{t_{j+1}}$ via $\yzl_{k,t_j} \ge \yzu_{k,t_{j+1}}$
as for relatively dense $t$ grids, neighbouring intervals should be able to overlap***

***Require $\yzu_{k,t_j} \ge \yzu_{k,t_{j+1}}$ and $\yzl_{k,t_j} \ge \yzl_{k,t_{j+1}}$.
If we know, e.g., pretty much what's going on for low $t$
but want to be much more cautious for high $t$,
then $\yktzl$ can drop to 0, but $\yktzu$ cannot increase.

***Better to express certainty of prior guess via $\nktz$ bounds?\\
With (strongly) differing $\nktz$ values, it can happen that, e.g., $\ynu_{k,t_j} < \ynu_{k,t_{j+1}}$!\\
E.g., when $\yzu_{k,t_j} = \yzu_{k,t_{j+1}}$ and $\yzl_{k,t_j} = \yzl_{k,t_{j+1}}$,
$s^k_t/n_k \in [\yzl_{k,t_j}, \yzu_{k,t_j}]$ (so no prior-data conflict) and no failures in $[t_j, t_{j+1}]$,\\
then with $\nzu_{k,t_j} < \nzu_{k,t_{j+1}}$ we have $\ynu_{k,t_j} < \ynu_{k,t_{j+1}}$,\\
 and with $\nzu_{k,t_j} > \nzu_{k,t_{j+1}}$ we have $\ynl_{k,t_j} > \ynl_{k,t_{j+1}}$.\\
(weighted average between interval and precise value: the lower $n$, the 'faster' the interval will shrink!)\\
So we need to require same $\nktz$ bounds for all $t$!
%(Prior guess for $p^k_t$ interpretation would say that $\yktz$ cannot increase,
%but does this need to hold for the prior guess range bound $\yktzu$?)***

***generally require $\yktzl > 0$ and $\yktzu < 1$ for all $t \not\in \big(\min(\vec{t}^k), \max(\vec{t}^k)\big)$
to get proper posteriors***

>>>>

\section{Nonparametric Bayesian Approach for Component Reliability}
\label{sec:nonparamapproach}

@@

For an informative prior the choice of $\nktzl$ is relevant for posterior inferences in case of prior-data conflict,
as then one of the bounds for $\ykt$ is attained for $\nktzl$.
In a near-noninformative setting, the choice of $\nktzl$ is instead not relevant,
as both $\yktnl$ and $\yktnu$ are obtained with $\nktzu$;
indeed $\yktzl > 0$ and $\yktzu < 1$ can be chosen such that
$\frac{s^k_t}{n_k} \in [\yktzl, \yktzu]$ for all $t \in \big(\min(\vec{t}^k), \max(\vec{t}^k)\big)$.
Obviously, one cannot have prior-data conflict in case of prior near ignorance.

@@

To obtain the lower and upper bound for the system reliability function $\Rsys(t)$,
we now need to minimise and maximise Equation~\eqref{eq:rsyswithsurvsign} over $\PtZi{1}, \ldots, \PtZi{K}$ for each $t$,
where the posterior predictive probabilities for $C^k_t$ are given by the Beta-Binomial pmf \eqref{eq:postpredCny}.
It turns out that we need to optimise over $\ntz{1}, \ldots, \ntz{K}$ only, %$\nktz$ ($k=1,\ldots,K$) only,
as lower and upper bounds for $\Rsys(t)$ are obtained for $\yktzl$ and $\yktzu$, respectively.

To see this,
remember that $\yktz$ gives the prior expected functioning probability for a component of type $k$ at time $t$.
From Equation~\eqref{eq:nyupdate} (right) we see that $\yktn$ is increasing in $\yktz$,
%the posterior parameter set bounds are the transformed prior parameter set bounds,
%i.e., $\yktzu$ corresponds to the upper bound of $\PktN$, for all $\nktz \in \PktZ$,
such that for each fixed $\nktz$, $\yktzu$ will maximise the posterior expected functioning probability $\yktn$.
%
Now, a higher posterior expected functioning probability means
higher probability for many components out of $m_k$ surviving,
i.e., probabilities for the larger $l_k$'s are higher, and in turn for the lower $l_k$'s are lower.
Indeed, the Beta-Binomial distribution is stochastically ordered for varying $\yktn$, i.e.,
its cmf \eqref{eq:postpredCnycmf} is decreasing in $\yktn$ for all $l_k$,
such that the cmfs for different values of $\yktn$ do not cross each other.
This means that for each fixed $\nktz$, the Beta-Binomial cmf \eqref{eq:postpredCnycmf} is minimized for $\yktzu$ in all $l_k$,
such that the probability weights given to high values of $l_k$ are maximised. %for $\yktzu$.
For coherent systems, $\Phi$ is non-decreasing in each of its arguments $l_1,\ldots,l_K$,
thus $\Rsys(t)$ is maximised by putting most probability weight on the highest $l_k$ values for each $k$,
which is in turn obtained for $\yktzu$, $k=1,\ldots,K$.
We therefore have
%
%for a certain $k$, a high probability on high $l_k$'s is never worse than a high probability on lower $l_k$'s,
%keeping all other component type expressions fixed.
%***Monotonicity in $\yz_{k,t}$, so we get, for each $t \in {\cal T}$,
\begin{align*}
\lefteqn{\lRsys(t \mid \vec{t}^1, \ldots, \vec{t}^K)}\\
 &= \min_{\PtZi{1}, \ldots, \PtZi{K}} \Rsys(t \mid \PtZi{1}, \ldots, \PtZi{K}, \vec{t}^1, \ldots, \vec{t}^K) \\
 &= \min_{\nz_{1,t},\ldots,\nz_{K,t}} 
    \sum_{l_1=0}^{m_1} \cdots \sum_{l_K=0}^{m_K} \Phi(l_1, \ldots, l_K)
                                                 \prod_{k=1}^K P(C^k_t = l_k \mid \yktzl, \nktz, s^k_t) \\
 &= \min_{\nz_{1,t},\ldots,\nz_{K,t}} 
    \sum_{l_1=0}^{m_1} \cdots \sum_{l_K=0}^{m_K} \Phi(l_1, \ldots, l_K) \times \\ & \hspace*{12ex}
    \prod_{k=1}^K {m_k \choose l_k} \frac{B(l_k + \nn_{k,t}\ynl_{k,t}, m_k - l_k + \nn_{k,t}(1-\ynl_{k,t}))}
                                         {B(\nn_{k,t}\ynl_{k,t}, \nn_{k,t}(1-\ynl_{k,t}))} \,.
\end{align*}
In the following, we will give some guidelines on how to choose the parameter sets $\PkZi{1}, \ldots, \PkZi{\tmax}$
which define the set of prior discrete reliability functions for components of type $k$.
We think this is much easier in terms of $\nz$ and $\yz$ than it would be in terms of $\alpha$ and $\beta$.
%***How to choose the sets of priors for the survival function of type $k$ components,
%i.e. how to choose $\PkZi{1}, \ldots \PkZi{\tmax}$?***

As mentioned in Section~\ref{sec:nonparamapproach}, for the actual functioning probabilities $\ptk$
it should hold that $p^k_{t_j} \ge p^k_{t_{j+1}}$. ***or $p^k_{t_i} \ge p^k_{t_i}$ for all $i < j$.
%***not necessary to ensure $p^k_{t_j} \ge p^k_{t_{j+1}}$ via $\yzl_{k,t_j} \ge \yzu_{k,t_{j+1}}$
%as for relatively dense $t$ grids, neighbouring intervals should be able to overlap***
%(Prior guess for $p^k_t$ interpretation would say that $\yktz$ cannot increase,
%but does this need to hold for the prior guess range bound $\yktzu$?)***
This naturally translates to conditions on the prior guess for $\ptk$,
such that we propose that 
$\yzu_{k,t_j} \ge \yzu_{k,t_{j+1}}$ and $\yzl_{k,t_j} \ge \yzl_{k,t_{j+1}}$ should hold.
Due to $s^k_t/n_k$ decrasing in $t$ and the weighted average property of the update step for $\ykt$,
this ensures that 
$\ynu_{k,t_j} \ge \ynu_{k,t_{j+1}}$ and $\ynl_{k,t_j} \ge \ynl_{k,t_{j+1}}$.
In a situation where we are quite sure about the functioning probability for low $t$,
but are unsure about what happens for larger $t$ and so want to be maximally cautious for those $t$,
then we can let $\yktzl$ drop to (almost) 0, but $\yktzu$ should not increase.
%***Require $\yzu_{k,t_j} \ge \yzu_{k,t_{j+1}}$ and $\yzl_{k,t_j} \ge \yzl_{k,t_{j+1}}$.
%If we know, e.g., pretty much what's going on for low $t$
%but want to be much more cautious for high $t$,
%then $\yktzl$ can drop to 0, but $\yktzu$ cannot increase.

We think it is not advisable to express certainty about expected functioning probabilities
with $\nktz$ bounds that vary strongly over $t$.
%***Better to express certainty of prior guess via $\nktz$ bounds?\\
With (strongly) differing $\nktz$ bounds, monotonicty for $\yktn$ bounds cannot be ensured.
%it can happen that, e.g., $\ynu_{k,t_j} < \ynu_{k,t_{j+1}}$!\\
E.g., when $\yzu_{k,t_j} = \yzu_{k,t_{j+1}}$, $\yzl_{k,t_j} = \yzl_{k,t_{j+1}}$,
and $s^k_{t_j}/n_k \in [\yzl_{k,t_j}, \yzu_{k,t_j}]$ (so there is no prior-data conflict),
and furthermore no failures happening in $[t_j, t_{j+1}]$, i.e., $s^k_{t_{j+1}}/n_k = s^k_{t_{j}}/n_k$, then
\begin{itemize}
\item for $\nzu_{k,t_j} < \nzu_{k,t_{j+1}}$ we get $\ynu_{k,t_j} < \ynu_{k,t_{j+1}}$, and
\item for $\nzu_{k,t_j} > \nzu_{k,t_{j+1}}$ we get $\ynl_{k,t_j} < \ynl_{k,t_{j+1}}$.
\end{itemize}
Again, this follows from the weighted average property of $\yktn$;
for constant $\yktz$, the lower $\nktz$ the closer $\yktn$ is to $s^k_t/n_k$.
% one can think of the update between interval and precise value: the lower $n$, the 'faster' the interval will shrink!)\\
Therefore, we propose to take the same $\nktz$ bounds for all $t$, or, at most, to have it change with $t$ gradually only.

%***generally require $\yktzl > 0$ and $\yktzu < 1$ for all $t \not\in \big(\min(\vec{t}^k), \max(\vec{t}^k)\big)$
%to get proper posteriors***
Generally, one should choose $\yktzl > 0$ and $\yktzu < 1$ for all $t \not\in \big(\min(\vec{t}^k), \max(\vec{t}^k)\big)$.
